Preparing your manuscript

### 2025-01-31 article/article.qmd
Figure 2 - first version

library(metR)
library(ggplot2)
library(data.table)
library(lubridate)
library(here)
library(cowplot)

@@

map_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "change") |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lon, lat)] |> 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_contour_fill(aes(z = mean_dt/365*100, fill = after_stat(level)),
                    breaks = c(-Inf, seq(-3, 3, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorbar(barheight = 0.5),
                                   labels = function(x) JumpBy(x, 2, fill = "")) +
  geom_sf(data = global_map, inherit.aes = FALSE, fill = NA, linewidth = 0.2) +
  # scale_y_latitude(labels = NULL) +
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL, fill = "%") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
        legend.key.width = grid::unit(1, 'null'),
        # axis.title.x=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), plot.margin = unit(c(0, 0 , 0 , 0), units = "mm"))

@@

>>>>

library(metR)
library(ggplot2)
library(data.table)
library(lubridate)
library(here)
library(cowplot)
library(ggside)

@@

map_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "change") |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lon, lat)] |> 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_contour_fill(aes(z = mean_dt/365*100, fill = after_stat(level)),
                    breaks = c(-Inf, seq(-3, 3, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorbar(barheight = 0.5),
                                   labels = function(x) JumpBy(x, 2, fill = "")) +
  geom_sf(data = global_map, inherit.aes = FALSE, fill = NA, linewidth = 0.2) +
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL, fill = "%") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
        legend.key.width = grid::unit(1, 'null'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), plot.margin = unit(c(0, 0 , 0 , 0), units = "mm"))

@@

(ref:serie)

```{r}
#| label: serie
#| fig-width: 10
#| fig-height: 7
#| fig.cap: (ref:serie)

serie <- purrr::map(Sys.glob(here("data/series/serie_freq_deltap97.5_*_100km.nc")), function(f) {

  meta <- unglue::unglue(basename(f), "serie_freq_{threshold}_{scenario}_100km.nc")
  ReadNetCDF(f, vars = c(mean_freq = "tasmax")) |> 
    _[, let(year = year(time),
            scenario = meta[[1]][["scenario"]])] |> 
    _[, let(time = NULL)]

}) |> rbindlist() |> 
  _[, let(scenario = factor(scenario, labels = c("Historical", "SSP5-8.5")))]

model_spread <- data |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50 - NH",
                        lat %between% c(38, 40) & land == "Sea", "40 - NH",
                        lat %between% c(28, 30) & land == "Sea", "30 - NH", 
                        lat %between% c(48, 50) & land == "Land", "50 - NH",
                        lat %between% c(38, 40) & land == "Land", "40 - NH",
                        lat %between% c(28, 30) & land == "Land", "30 - NH",
                        lat %between% c(-50, -48) & land == "Land", "50 - SH",
                        lat %between% c(-40, -38) & land == "Land", "40 - SH",
                        lat %between% c(-30, -28) & land == "Land", "30 - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50 - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40 - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30 - SH"))] |>
  _[!is.na(serie) & var == "deltap97.5" & model %in% models_in_gadi] |>
    tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[, .(mean_freq = mean(ssp585)), by = .(lat, hemisphere, land, model)] |> 
  _[, let(scenario = "SSP5-8.5")]

serie_regions <- serie |>
  _[!(year == 2014 & scenario != "historical")] |>
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50 - NH",
                        lat %between% c(38, 40) & land == "Sea", "40 - NH",
                        lat %between% c(28, 30) & land == "Sea", "30 - NH", 
                        lat %between% c(48, 50) & land == "Land", "50 - NH",
                        lat %between% c(38, 40) & land == "Land", "40 - NH",
                        lat %between% c(28, 30) & land == "Land", "30 - NH",
                        lat %between% c(-50, -48) & land == "Land", "50 - SH",
                        lat %between% c(-40, -38) & land == "Land", "40 - SH",
                        lat %between% c(-30, -28) & land == "Land", "30 - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50 - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40 - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30 - SH"))] |>
  _[!is.na(serie)] |> 
  _[, .(mean_freq = mean(mean_freq)), by = .(land, serie, year, scenario)]

serie_regions |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  # _[, let(lat = forcats::fct_relevel(lat, "55N", after = 0))] |> 
  ggplot(aes(year, mean_freq, color = lat, group = interaction(scenario, lat))) +
  geom_line(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_ysideboxplot(data = model_spread, aes(x = lat), orientation = "x") +
  scale_color_manual(values = c("#004C73", "#CC3311", "#EE7733")) +
  facet_grid(land~hemisphere, scale = "free") +
  labs(x = "Year", y = "Frequency (days/year)", color = "Latitude") +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text = element_text(size = 12),
        ggside.axis.text.x = element_blank())

```{r}
serie |>
  _[!(year == 2014 & scenario != "historical") & lat < 0] |>
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(band_lat = cut_width(lat, 5))] |> 
  _[, let(mean_freq = mean(mean_freq)), by = .(land, band_lat, year, scenario)] |> 
  _[, FitLm(mean_freq, year),  by = .(land, band_lat, scenario)] |> 
  _[term == "year"] |> 
  ggplot(aes(scenario, estimate)) +
  geom_point(aes(color = band_lat)) +
  geom_path(aes(color = band_lat)) +
  facet_wrap(~land)
```
